---
title: "Exploratory Data Analysis of Breweries in United States"
author: "Garrett Drake and Eny Ntuk"
date: "2022-10-16"
output: html_document

---

## Objective
### This studies is to analyze the ABV (Alcohol By Volume) versus the IBU (International Bitterness Unit) of 2410 US craft beers made by 558 US breweries and to also show any relationship between the ABV and IBU of beers.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries}
# load libraries
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(knitr)
library(magrittr)
library(kableExtra)
library(class)
library(caret)
library(e1071)
library(RCurl)
library(stringr)
library(psych)
```

`````{r}
# load data
beer <- read.csv("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%208%20and%209%20Case%20Study%201/Beers.csv", header = T)

breweries <- read.csv("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%208%20and%209%20Case%20Study%201/Breweries.csv", header = T)
```

# Problem 1
```{r Problem 1 Breweries per State}
######## Problem 1 ----- How many breweries are in each state #####
breweries$State <- factor(breweries$State) # put into factor so can count
levels(breweries$State) # check levels
num_breweries <- breweries %>% count(State) # get total count of breweries by state into new DF
num_breweries <- rename(num_breweries, total_number = n) # rename column to total_number
table(breweries$State)
```

# Problem 1 Printout
## With plot of breweries by state in order from least to greatest
```{r Problem 1 printout}
kable(num_breweries, caption = "Problem 1 Breweries per state")
```


```{r plot breweries by state in order from least to greatest }
# plot breweries by state in order from least to greatest
num_breweries %>% ggplot(aes(x=reorder(State,total_number), y = total_number)) +
  geom_bar(aes(fill = State), position = "dodge", stat = "identity") +
  geom_text(aes(label=total_number), position=position_dodge(width=0.9), size=2.0, vjust=-0.35) +
  theme(axis.text.x=element_text(size=rel(0.8), angle=90)) +
  ggtitle("Total Breweries by State") +
  labs(x="State",y="Brewery Count")

```



## Problem 2 Merege data sets and provide first and last 6 rows
```{r Problem 2 Merge Data sets}
####### Problem 2 Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.) ########
big_beer <- full_join(beer, breweries, by = c("Brewery_id" = "Brew_ID" ))
big_beer <- rename(big_beer, Beer = Name.x, Brewery = Name.y, Beer_Type = Style)
```

### First 6 rows of merged data set
```{r}
head(big_beer, n =6)
```

### Last 6 rows of merged data set
```{r}
tail(big_beer, n = 6)
```

## Problem 3 Address missing values
### There are 62 missing values in ABV and 1005 in IBU
```{r Problem 3 - missing values}
######### Problem 3 - address missing values ######
colSums(is.na(breweries)) # no missing values in the breweries data set
colSums(is.na(beer)) # ABV has 62 missing values, IBU has 1005 missing values
colSums(is.na(big_beer)) # sanity check for merged data ABV has 62 missing values, IBU has 1005 missing values
```


```{r Problem 3 missing values with kable}
# Used kable Extra library to make document more presentable
colSums(is.na(big_beer)) %>%  
  kable("html", caption = "Problem 3 Missing Values in Each Variable") %>%
  kable_styling()
```

## Problem 4

```{r Problem 4 Data Frame with median ABV and IBU by State}
######### Problem 4	Compute median ABV and IBU for each state. Plot a bar chart to compare. ########
state_med_ABV_IBU <- big_beer %>% group_by(State) %>% summarise(Median_ABV = median(ABV, na.rm = TRUE), Median_IBU = median(IBU, na.rm = TRUE)) # create data set with medians ABV and IBU for each state
```

### Table of Median ABV and IBU by State
```{r}
kable(state_med_ABV_IBU, caption = "Table of Median ABV and IBU by State")
```

### Barchart of Median ABV by State ordered from least to greatest
```{r plot of median ABV by state with color and ordered from greatest to least}
# create bar chart for median ABV by state with color and ordered from least to greatest
state_med_ABV_IBU %>% ggplot(aes(x=reorder(State,Median_ABV), y = Median_ABV)) +
  geom_bar(aes(fill = State), stat = "identity") +
  geom_text(aes(label=Median_ABV), position=position_stack(vjust = 1.05), size=2.0, angle = 90) +
  theme(axis.text.x=element_text(size=rel(0.8), angle=90)) +
  ggtitle("Median ABV by State") +
  labs(x="State",y="ABV")
```

### Barchart of Median IBU by State ordered from least to greatest
```{r plot of median IBU by state with more color order by least to greatest}
#create bar chart for median IBU by state with more color order by least to greatest
state_med_ABV_IBU %>% ggplot(aes(x=reorder(State,Median_IBU), y = Median_IBU)) +
  geom_bar(aes(fill = State), stat = "identity") +
  geom_text(aes(label=Median_IBU), position=position_stack(vjust = 1.05), size=2.0, angle = 90) +
  theme(axis.text.x=element_text(size=rel(0.8), angle=90)) +
  ggtitle("Median IBU by State") +
  labs(x="State",y="IBU")
```

## Problem 5
Colorado has the maximum ABV beer named “Lee Hill Series Vol. 5 - Belgian Style Quadrupel Ale” with an ABV of .128

Oregon has the maximum IBU beer named “Bitter Bitch Imperial IPA” with an IBU of 138
```{r Problem 5 - State with Max ABV and IBU}
########### Problem 5 Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer? ##########
which.max(big_beer$ABV) # gives number of which state has max ABV, which is number 2279
big_beer$State[2279] # 2279 is CO, so Colorado has max ABV
max(big_beer$ABV, na.rm = T) # highest ABV is .128
big_beer$Beer[2279] # Name of highest ABV Beer is "Lee Hill Series Vol. 5 - Belgian Style Quadrupel Ale"
big_beer$ABV[2279] # sanity check, all lings up

which.max(big_beer$IBU) # gives number of which state has max IBU, which is number 148
big_beer$State[148] # 148 is OR, so Oregon has max IBU
max(big_beer$IBU, na.rm = T) # highest IBU is 138
big_beer$Beer[148] # Name of highest IBU is "Bitter Bitch Imperial IPA"
big_beer$IBU[148] # sanity check, all lings up
```

## Problem 6
Summary Statistics

The minimum ABV is .001 and maximum is .128  
The median ABV is .056  
The mean ABV is .0598, standard deviation is .014, and there were 62 missing values  
50% of the distribution of beer ABV is between .05 and .067

```{r Problem 6 summary statistics}
######## Problem 6.	Comment on the summary statistics and distribution of the ABV variable. #########
summary(big_beer$ABV) # get summary statistics
sd(big_beer$ABV, na.rm = TRUE) # get SD of ABV
```

## Boxplot of beer ABV
```{r box plot of beers}
#box plot of beers
big_beer %>% ggplot(aes(x=ABV)) +
  geom_boxplot() +
  theme_economist() + 
  scale_color_economist() +
  ggtitle("Boxplot of ABV")
```

## Histogram of beer ABV
```{r Histogram of beers}
# Histogram of beers
big_beer %>% ggplot(aes(x=ABV)) +
  geom_histogram( colour="blue", fill="white") +
  ggtitle("Distribution of beers by ABV")
```


## Problem 7
```{r Problem 7 Relationship between IBU and ABV}
################ Problem 7.	Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer. #########

big_beer %>% ggplot(aes(x = IBU, y = ABV)) +
  geom_point() +
  geom_smooth(method=lm) + # add linear regression line
  theme_economist() + 
  scale_color_economist()+
  theme(axis.text.x=element_text(size=rel(1.0)))+
  ggtitle("Correlation between IBU and ABV ") +
  labs(x="IBU",y="ABV")
```

### There is evidence to suggest that a medium positive relationship between IBU and ABV exists,i.e. generally as IBU increases so does ABV and vice versa. A linear estimate of this relationship is shown by the blue line. Correlation coefficient r = .671, representing medium correlation

``` {r correlation coefficient}
cor(big_beer$IBU, big_beer$ABV, use = 'pairwise.complete.obs', method = "pearson")
```


# Problem 8
## Build Hypertuned Knn model
```{r Hypertuned Knn model}
beerKnn <- na.omit(big_beer) %>% select(ABV, IBU, Beer_Type) %>%  filter(grepl("IPA|Ale", Beer_Type, ignore.case = TRUE)) # create data frame to do Knn test with ABV, IBU and Beer types (IPA or Ale)

beerKnn$ipa_ale <- factor(ifelse(grepl("IPA", beerKnn$Beer_Type, ignore.case = TRUE), "IPA", "Ale"))
# Build Hypertuned model to find best K
set.seed(317)
iterations <-  100 # 200 iterations takes like 1.5 mins to run, consider smaller, tried 60 and much faster with minimal change in results
numks <-  60 # gives back 60 iterations of k values
splitPerc <-  .75 # 75 percent split
masterAcc <-  matrix(nrow = iterations, ncol = numks)
for(j in 1:iterations)
{
  trainIndices = sample(1:dim(beerKnn)[1],round(splitPerc * dim(beerKnn)[1]))
  train <-  beerKnn[trainIndices , ]
  test <-  beerKnn[-trainIndices , ]
  for(i in 1:numks)
  {
    classifications = knn(train[,c("ABV", "IBU")],test[,c("ABV", "IBU")],train$ipa_ale, prob = TRUE, k = i)
    table(classifications,test$ipa_ale)
    CM = confusionMatrix(table(classifications,test$ipa_ale))
    masterAcc[j,i] = CM$overall[1]
  }
  
}
MeanAcc <- colMeans(masterAcc)
which.max(MeanAcc) # which K gives max percentage
max(MeanAcc)
```

```{r Plot of model}
plot(seq(1,numks,1),MeanAcc, type = "l", xlab = "Number of Ks", ylab = "Mean Accuracy", main = "k-NN Classification Accuracy for IPA and IBV Beers", col = "blue")
abline(v = 5, col = "red")
abline(h = .8607884, col = "red")
```

# Addition things
T test for ABV and IBU
```{r t tests for abv and IBU}
t.test(ABV ~ ipa_ale, data = beerKnn)
t.test(IBU ~ ipa_ale, data = beerKnn)
```

### Session Info:
```{r}
sessionInfo()
```


